---
output: github_document
---

<!-- Paper_results_2.md is generated from paper_results_2.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/results2-",
  dev = "png",
  out.width = "80%"
)
```

# bremla

This file includes the data and code needed to reproduce the results for

Myrvoll-Nilsen, E., Riechers, K. & Boers, N. (202x). Tie-point paper (title TBD)

# Myrvoll-Nilsen et al. (202x)

The data used is the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm', downloaded from [Centre for Ice and Climate](iceandclimate.nbi.ku.dk) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [Rasmussen et al. (2014)](https://www.sciencedirect.com/science/article/pii/S0277379114003485). These are included in the package and can be imported as follows. 
```{r install, eval=FALSE}
#library(bremla)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")

age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
proxy = NGRIP_5cm$d18O
n=length(diff(age))

data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
events=list(locations = events_rasmussen$depth,
            locations_unit="depth",degree=1)

plot(depth,proxy,xlab="Age (yb2k)",ylab=expression(paste(d^18,"O (permil)")),type="l",
     xlim=rev(range(depth)))+abline(v=events$locations,col="gray",lwd=0.7)
formula = dy~-1+depth2+proxy



```


We also use tie-points described by probability distributions obtained from Adolphi et al. (2018) and Muscheler et al. (2020). These can be imported as follows.

```{r adolphi, eval=TRUE}

adolphipdfs = adolphiloader(tieshifts=c(11050,12050,13050,22050,42050),plotdens = TRUE,
                            x.ref=c(11050,12050,13050,22050,42050))

```

The model can be fitted using the \texttt{bremla} function, which estimates the posterior marginal distributions of the hyperparameters and also performs simulations of chronologies.


```{r hyperpars, eval=FALSE}
nsims=10000

results = bremla(formula,data,reference.label="GICC05",
                nsims=nsims,
                events=events,
                control.fit=list(noise="ar1"),
                control.sim=list(synchronized=FALSE) ,
                print.progress=TRUE
                )

{
  
  par(mfrow=c(1,2))
  plot(results$fitting$inla$hyperparameters$posteriors$sigma_epsilon,type="l",xlab=expression(paste(sigma[epsilon])),ylab="Density")
  
  abline(v=results$fitting$inla$hyperparameters$results$sigma_epsilon$mean)
  abline(v=c(results$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.025,
             results$fitting$inla$hyperparameters$results$sigma_epsilon$quant0.975),
             col="red")
         
  plot(results$fitting$inla$hyperparameters$posteriors$phi,type="l",xlab=expression(phi),ylab="Density")
  abline(v=results$fitting$inla$hyperparameters$results$phi$mean)
  abline(v=c(results$fitting$inla$hyperparameters$results$phi$quant0.025,
             results$fitting$inla$hyperparameters$results$phi$quant0.975),
             col="red")
 par(mfrow=c(1,1)) 
}

```

The fitted model to the layer increments and the simulated chronologies (as compared to the GICC05 time scale) can be plotted as follows.

```{r unconstrained}
{
  library(ggplot2)
  
  dsims = colDiffs(rbind(matrix(rep(results$initial$age,ncol(results$simulation$age)),nrow=1),results$simulation$age))
  dmeans = rowMeans(dsims)
  dsd = rowSds(dsims)
ggd = data.frame(y = results$data$dy,depth=results$data$depth,
                   mean=dmeans,
                   lower = dmeans-1.96*dsd,
                   upper = dmeans+1.96*dsd)
  ggp = ggplot(data=ggd,aes(x=depth)) + theme_bw()+ #xlim(rev(range(results$data$depth)))+
    geom_line(aes(y=y),col="gray")+
    geom_line(aes(y=mean),col="blue",size=0.3)+
    geom_ribbon(aes(ymin=lower,ymax=upper),color="red",alpha=0,size=0.05)+
    xlab("Depth (m)")+ylab("Layers per 5cm")
 print(ggp) 
 
 #ggsave("fitdiff-4800x2000.eps", device=cairo_ps,width=4800,height=2000,units="px",dpi=500,limitsize=FALSE)
}

{
  gicc05=results$original.chron$age[1+1:n]
  ggd = data.frame(depth=results$data$depth,
                   mean=results$simulation$summary$mean-gicc05,
                   lower=results$simulation$summary$lower-gicc05,
                   upper=results$simulation$summary$upper-gicc05,
                   zeros = numeric(n))
  ggp = ggplot(data=ggd,aes(x=depth)) + theme_bw()+
    #geom_line(aes(y=zeros),color="gray",linetype="dashed")+
    geom_line(aes(y=mean),col="blue")+
    geom_ribbon(aes(ymin=lower,ymax=upper),fill="red",col="red",alpha=0.3)+
    xlab("Depth (m)") + ylab("Simulated time scale - GICC05 (years)")
  print(ggp)
  
  #ggsave("unsync-4800x2000.eps", device=cairo_ps,width=4800,height=2000,units="px",dpi=500,limitsize=FALSE)
}
```


## Fixed tie-points


In the paper we demonstrate our model first using the Adolphi tie-points fixed to be equal to the mode of the probability distribution. We omit the first tie-point since it falls outside of our considered interval. The constrained chronologies can be computed as follows.

```{r fixed, eval=TRUE}



fixedsims = cbind(#rep(inla.mmarginal(adolphipdfs$tie1),nsims),
                  rep(inla.mmarginal(adolphipdfs$tie2),nsims),
                  rep(inla.mmarginal(adolphipdfs$tie3),nsims),
                  rep(inla.mmarginal(adolphipdfs$tie4),nsims),
                  rep(inla.mmarginal(adolphipdfs$tie5),nsims)
                  )
synchronization = list(samples = fixedsims,nsims=nsims,
                       locations=c(#1492.50,
                                   1502.80,1532.70,1767.25,2132.40),
                       locations_unit="depth")

results_fixed = bremla(formula,data,reference.label="GICC05",
                nsims=nsims,
                events=events,
                synchronization=synchronization,
                control.fit=list(noise="ar1"),
                control.sim=list(synchronized=TRUE) ,
                print.progress=TRUE
                )

{
  gicc05=results_fixed$original.chron$age[1+1:n]
  ggd = data.frame(depth=results_fixed$data$depth,
                   mean=results_fixed$simulation$summary_sync$mean-gicc05,
                   lower=results_fixed$simulation$summary_sync$lower-gicc05,
                   upper=results_fixed$simulation$summary_sync$upper-gicc05,
                   zeros = numeric(n))
  ggp = ggplot(data=ggd,aes(x=depth)) + theme_bw()+
    #geom_line(aes(y=zeros),color="gray",linetype="dashed")+
    geom_line(aes(y=mean),col="blue")+
    geom_ribbon(aes(ymin=lower,ymax=upper),fill="red",col="red",alpha=0.3)+
    xlab("Depth (m)") + ylab("Simulated time scale - GICC05 (years)")
  print(ggp)
  
  #ggsave("fixed-4800x2000.eps", device=cairo_ps,width=4800,height=2000,units="px",dpi=500,limitsize=FALSE)
}


```

Synchronizing our chronologies with the Adolphi tie-points where the uncertainties is incorporated cam be done using Monte Carlo simulations. All of this can be done using the bremla function.

```{r sync, eval = TRUE}
library(ggplot2)
synchronization = list(method="adolphi",nsims=nsims,
                       locations=c(#1492.50,
                                   1502.80,1532.70,1767.25,2132.40),
                       locations_unit="depth")

results_sync = bremla(formula,data,reference.label="GICC05",
                nsims=nsims,
                events=events,
                synchronization=synchronization,
                control.fit=list(noise="ar1"),
                control.sim=list(synchronized=TRUE,summary=list(CI.type="hpd")) ,
                print.progress=TRUE
                )

tiepointranges = as.data.frame(matrix(NA,nrow=5,ncol=3))

gicc05=results_sync$original.chron$age[1+1:n]
gicc05tie = gicc05[results_sync$tie_points$locations_indexes]

colnames(tiepointranges)=c("lower","upper","depth")
for(i in 2:5){
  tiepointranges[i,1:2] = inla.hpdmarginal(0.95,adolphipdfs[[paste0("tie",i)]])-gicc05tie[i]
}
tiepointranges[,3] = depth[results_sync$tie_points$locations_indexes]

{
  
  ggd = data.frame(depth=results_sync$data$depth,
                   mean=results_sync$simulation$summary_sync$mean-gicc05,
                   lower=results_sync$simulation$summary_sync$lower-gicc05,
                   upper=results_sync$simulation$summary_sync$upper-gicc05,
                   zeros = numeric(n))
  ggp = ggplot(data=ggd,aes(x=depth)) + theme_bw()+
    #geom_line(aes(y=zeros),color="gray",linetype="dashed")+
    geom_line(aes(y=mean),col="blue")+
    geom_ribbon(aes(ymin=lower,ymax=upper),fill="red",col="red",alpha=0.3)+
    xlab("Depth (m)") + ylab("Simulated time scale - GICC05 (years)") +
    geom_linerange(data=tiepointranges,aes(x=depth,ymin=lower,ymax=upper),color="magenta")
  print(ggp)
  
  #ggsave("sync-4800x2000.eps", device=cairo_ps,width=4800,height=2000,units="px",dpi=500,limitsize=FALSE)
}

```


# Applications




## Abrupt warmings

The onset of abrupt warmings is dated in two stages that are combined to express the complete dating uncertainty, pertaining to the uncertainty originating from both the age-depth relationship as well as estimating the depth in the record which represent the onset of the transition. The latter is estimated using a linear ramp model as described in Erhardt et al. (2019). The linear ramp fit can be performed and plotted as follows.



```{r events_individual, eval=FALSE}

eventnumber=13 #number between 1 and 29. specifies which transition to consider

#load data window and specifics to transition
lowerints = which.index(event_intervals$depth_int_lower.m, depth[2:length(depth)])
upperints = which.index(event_intervals$depth_int_upper.m, depth[2:length(depth)])
depth.reference = event_intervals$NGRIP_depth_m[eventnumber]
age.reference = event_intervals$GICC_age.yb2k[eventnumber]
interval = lowerints[eventnumber]:upperints[eventnumber]

transitionlabel = str_sub(event_intervals$onsetlabel[eventnumber],
                      str_locate(event_intervals$onsetlabel[eventnumber],"GI")[1])


control.linramp=list(proxy=proxy,
                     interval=interval,
                     interval.unit="index",
                     depth.reference=event_intervals$NGRIP_depth_m[eventnumber]
                     )

results_event = linrampfitter(results_sync,control.linramp)

control.transition_dating = list(label="Simulated transition")
results_event = events_depth_to_age(results_event,
                                    control.transition_dating=control.transition_dating)
{
  layout(mat=matrix(c(1,2,1,3),nrow=2))
  par(mar=c(5,4,4,2)+0.1)
  
  plot(depth[interval],proxy[interval],type="l",col="gray",
     main=paste0("(a) ",transitionlabel," linear ramp fit"),
     xlab="Depth (m)", ylab=expression(paste(delta^18,"O (permil)")),
     xlim = rev(range(depth[interval]))
     )
lines(results_event$linramp$data$x,results_event$linramp$linrampfit$mean,lwd=1.5)
lines(results_event$linramp$data$x,results_event$linramp$linrampfit$q0.025,lwd=1.5,col="red")
lines(results_event$linramp$data$x,results_event$linramp$linrampfit$q0.975,lwd=1.5,col="red")
abline(v=depth.reference,lty=3,lwd=0.7,col="black")

ybottom = min(results_event$linramp$data$y)
ytop = min(results_event$linramp$linrampfit$q0.025)
margt0=results_event$linramp$param$t0$marg.t0
normt0.y = margt0[,2]/diff(range(margt0[,2]))*(ytop-ybottom)-min(margt0[,2])+ybottom
lines(x=margt0[,1],y=normt0.y,col="blue",lwd=2)
ybottom = min(results_event$linramp$data$y)
ytop = min(results_event$linramp$linrampfit$q0.025)
margt1 = results_event$linramp$param$t1$marginal
normt1.y = margt1[,2]/diff(range(margt1[,2]))*(ytop-ybottom)-min(margt1[,2])+ybottom
lines(x=margt1[,1],y=normt1.y,col="blue",lwd=2,lty=3)


plot(results_event$linramp$param$t0$marg.t0,type="l",xlab=expression(paste("Onset depth (m)")),ylab="Density",
     main="(b) Marginal posterior of onset depth")#, 
     #xlim=rev(range(depth.reference,results_event$linramp$param$t0$marg.t0[,1])))
#abline(v=depth.reference,lty=3)
abline(v=results_event$linramp$param$t0$mean,lwd=1.5)
abline(v=results_event$linramp$param$t0$q0.025,col="red")
abline(v=results_event$linramp$param$t0$q0.975,col="red")

plot(results_event$linramp$param$t1$marginal,type="l",xlab=expression(paste("End point depth (m)")),ylab="Density",
     main="(c) Marginal posterior of end point depth")#, 
     #xlim=rev(range(depth.reference,results_event$linramp$param$t0$marg.t0[,1])))
#abline(v=depth.reference,lty=3)
abline(v=results_event$linramp$param$t1$mean,lwd=1.5)
abline(v=results_event$linramp$param$t1$q0.025,col="red")
abline(v=results_event$linramp$param$t1$q0.975,col="red")


par(mfrow=c(1,1),mar=c(5,4,4,2)+0.1)

hist(results_event$event_dating$samples,main=paste0(transitionlabel," Onset age"),
     xlab="Onset age (yr b2k)",col="orange",xlim=rev(range(results_event$event_dating$samples)), breaks=40)
abline(v=age.reference,lty=3)
abline(v=results_event$event_dating$mean,lwd=2)
abline(v=results_event$event_dating$q0.025,lwd=2,col="red")
abline(v=results_event$event_dating$q0.975,lwd=2,col="red")

}

```


We want to compare the dating onsets with those given by Rasmussen et al. (2014), Capron (2021) and Buizert et al. (2015). We start by importing those values

```{r events_import, eval=FALSE}
library(readxl)
rasmussen_depths = event_intervals$NGRIP_depth_m
rasmussen_ages = event_intervals$GICC_age.yb2k #GISevents$`Age (a b2k)`

buizert_onsets = read_excel("inst/Examples/Paper_results/data/Buizert_onsets.xlsx",
                            col_types=c("text","numeric","numeric","numeric","numeric"))
buizert_depths = buizert_onsets$Depth; buizert_ages = buizert_onsets$Age
#buizert_depths[c(2:6,8:9,12:13,15:17,19:21)]
buizert_ages = buizert_ages[c(2:6,8:9,12:13,15:17,19:21)]

capron_onsets_NGRIP_d18O = read_excel("inst/Examples/Paper_results/data/Capron_onsets.xls",
                           sheet="d18O",n_max=25)
capron_ages_NGRIP_d18O = capron_onsets_NGRIP_d18O$`t1 (50%)`

capron_onsets_NGRIP_dust = read_excel("inst/Examples/Paper_results/data/Capron_onsets.xls",
                                sheet="Ca2+",n_max=24)
capron_ages_NGRIP_dust = capron_onsets_NGRIP_dust$`t1 (50%)`

capind = numeric(29); buiind = numeric(29)

capind = c(NA,2,3, 4,5,6,NA,NA,7,8,NA,9,10,11,NA,NA,NA,NA,NA,12,13,14,NA,NA,15,NA,NA,16,17)
buiind = c(NA,2,NA,3,4,6,NA,NA,8,9,NA,11,12,13,NA,NA,NA,NA,NA,15,16,17,NA,NA,19,NA,NA,20,21)
```

We then estimate the dating uncertainty for all 29 events
```{r}
agesimmatrix_d18O = matrix(NA,29,nsims)
depthlist_d18O = c()
do.plot.rampfit = TRUE
do.plot.agehist = FALSE
depthstats = as.data.frame(matrix(NA,4,29))
colnames(depthstats) = c("true", "mean", "CIL","CIU","dustmean","dustCIL","dustCIU")
agestats = as.data.frame(matrix(NA,4,29))
colnames(agestats) = c("true", "mean", "CIL","CIU","dustmean","dustCIL","dustCIU")

steplengths = rep(0.01,29)
steplengths[21]=0.001 #sometimes changing the steplength slightly might improve convergence (default is 0.01)

for(eventnumber in 1:29){
  transitionlabel = str_sub(event_intervals$onsetlabel[eventnumber],
                      str_locate(event_intervals$onsetlabel[eventnumber],"GI")[1])
  lowerints = which.index(event_intervals$depth_int_lower.m, results_sync$data$depth)
  upperints = which.index(event_intervals$depth_int_upper.m, results_sync$data$depth)
  interval = lowerints[eventnumber]:upperints[eventnumber]
  
  depth.reference = event_intervals$NGRIP_depth_m[eventnumber]
  age.reference = event_intervals$GICC_age.yb2k[eventnumber]
  
  control.linramp = list(proxy=proxy,interval=interval,interval.unit="index",
                         depth.ref=depth.reference, 
                         label=paste0(eventnumber," (d18O):", event_intervals[eventnumber,2])
                         )
  
  temp_d18O = linrampfitter(results_sync,control.linramp)
  if(do.plot.rampfit){
    {plot(depth[interval],proxy[interval],type="l",col="gray",
          xlim=rev(range(depth[interval])), xlab="Depth (m)",ylab=expression(paste(delta^18,"O (permil)")),
          main=paste0(eventnumber,": ", transitionlabel))
      lines(temp_d18O$linramp$data$x,temp_d18O$linramp$linrampfit$mean)
      lines(temp_d18O$linramp$data$x,temp_d18O$linramp$linrampfit$q0.025,col="red")
      lines(temp_d18O$linramp$data$x,temp_d18O$linramp$linrampfit$q0.975,col="red")
      abline(v=rasmussen_depths[eventnumber],lty=3)
      
      ybottom = min(temp_d18O$linramp$data$y)
      ytop = min(temp_d18O$linramp$linrampfit$q0.025)
      margt0=temp_d18O$linramp$param$t0$marg.t0
      normt0.y = margt0[,2]/diff(range(margt0[,2]))*(ytop-ybottom)-min(margt0[,2])+ybottom
      lines(x=margt0[,1],y=normt0.y,col="blue",lwd=2)
      ybottom = min(temp_d18O$linramp$data$y)
      ytop = min(temp_d18O$linramp$linrampfit$q0.025)
      margt1 = temp_d18O$linramp$param$t1$marginal
      normt1.y = margt1[,2]/diff(range(margt1[,2]))*(ytop-ybottom)-min(margt1[,2])+ybottom
      lines(x=margt1[,1],y=normt1.y,col="blue",lwd=2,lty=3)
    }
  }
  
  
  control.transition_dating=list(label="d18O")
  agetemp_d18O = events_depth_to_age(temp_d18O,control.transition_dating)
  
  
  agesimmatrix_d18O[eventnumber,] = agetemp_d18O$event_dating$samples 
  if(do.plot.agehist){
    hist(agesimmatrix_d18O[eventnumber,],freq=0,breaks=50,col="orange",main=paste0("Onset age: ",transitionlabel),xlab="Age (yb2k)")
    abline(v=age.reference,col="blue")
  }
  
  #Store summary statistics
  depthstats[1,eventnumber] = event_intervals$NGRIP_depth_m[eventnumber]
  depthstats[2,eventnumber] = agetemp_d18O$linramp$param$t0$mean
  depthstats[3,eventnumber] = agetemp_d18O$linramp$param$t0$q0.025
  depthstats[4,eventnumber] = agetemp_d18O$linramp$param$t0$q0.975
  
  densage = density(agesimmatrix_d18O[eventnumber,]); densage = data.frame(x=densage$x,y=densage$y)
  
  zage = inla.zmarginal(densage,silent=TRUE)
  
  agestats[1,eventnumber] = event_intervals$GICC_age.yb2k[eventnumber]
  agestats[2,eventnumber] = mean(agesimmatrix_d18O[eventnumber])
  agestats[3,eventnumber] = zage$quant0.025
  agestats[4,eventnumber] = zage$quant0.975
  
}

```


```{r}

{
par(mfrow=c(5,6),mar=c(2,2,1.5,1))
for(i in 1:29){
  dens = density(agesimmatrix_d18O[i,])
  dens = data.frame(x=dens$x,y=dens$y)#/max(dens2$y))
  xlim = range(dens$x,event_intervals$GICC_age.yb2k[i]);ylim = range(dens$y)
  
  if(!is.na(buiind[i])) xlim=range(xlim,buizert_onsets$Age[buiind[i]])
  if(!is.na(capind[i])) xlim=range(xlim,capron_ages_NGRIP_d18O[capind[i]])
  #par(mar=c(2,2,1.5,1))
  
  
  plot(dens,type="l",col=1,xlab="Onset year (b2k)",ylab="Density",
       main=str_sub(event_intervals[i,2],str_locate(event_intervals[i,2],"GI-")[1]),
       xlim=xlim,ylim=ylim)
  #Axis(side=1,labels=FALSE,at=dens1$x,tck=0.1)
  
  #lines(dens,type="l",col="gray")
  if(FALSE){#95% credible intervals
    abline(v=mean(agesimmatrix_dust[i,]),col=1)
    zdens1 = inla.zmarginal(dens1,silent = TRUE)
    abline(v=c(zdens1$quant0.025,zdens1$quant0.975),col=1,lty=3,lwd=0.8)
    abline(v=mean(agesimmatrix_d18O[i,]),col=2)
    zdens2 = inla.zmarginal(dens2,silent = TRUE)
    abline(v=c(zdens2$quant0.025,zdens2$quant0.975),col=2,lty=3,lwd=0.8)
  }
  
  abline(v=event_intervals$GICC_age.yb2k[i],col="blue")  
  if(!is.na(buiind[i])){
    abline(v=buizert_onsets$Age[buiind[i]],col="green")
    #abline(v=buizert_ages[buiind[i]],col="green")
  }
  
  
  # abline(v=capron_ages_NGRIP_d18O,col="orange")
  # abline(v=capron_ages_NGRIP_dust,col="orange",lty=3)
  if(!is.na(capind[i])){
    abline(v=capron_ages_NGRIP_d18O[capind[i]],col="red")
  }
  
  
}
par(mar=c(0,0,0,0));plot(-1,axes=FALSE,xlab="",ylab="",xlim=c(0,1),ylim=c(0,1))
legend(0,1,legend=c("Onset posterior distribution",#"Ca2+ onset posterior",
                    "Rasmussen onset","Buizert onset",
                    "Capron onset"),#"Capron Ca2+ onset"),
       col=c("black", "blue", "green", "red"),
       lty=c(1,1,1,1,1,1), cex=1,bty="n")

par(mfrow=c(1,1),mar=c(5,4,4,2)+0.1)

}

rownames=c()
for(i in 1:29){
  rownames=c(rownames,str_sub(event_intervals[i,2],str_locate(event_intervals[i,2],"GI-")[1]))
}

#colnames = colnames(depthstats)[1:4]
#colnames = c("name","d18O_mean","d18O_q0.025","d18O_q0.975")
depthresults = t(depthstats)[,2:4]
colnames(depthresults)=c("Z* mean", "Z* q0.025","Z* q0.975")

ageresults = t(agestats[2:4,])
colnames(ageresults)=c("Y* mean", "Y* q0.025","Y* q0.975")
allresults = cbind(depthresults,ageresults)
rownames(allresults)=rownames

latexstr = "Event & Z^* mean (m) & Z^* 95\\% CI (m) & Y^* mean (yb2k) & Y^* 95\\% CI (yb2k) \\\\ \n"
for(i in 1:29){
  latexstr = paste0(latexstr,rownames[i], "& ", format(round(allresults[i,1],2),nsmall=2)," & (", format(round(allresults[i,2],2),nsmall=2), ",
         ",format(round(allresults[i,3],2),nsmall=2),") & ", format(round(allresults[i,4],2),nsmall=2)," & (",format(round(allresults[i,5],2),nsmall=2),", ",format(round(allresults[i,6],2),nsmall=2),") \\\\ \n"
         )
}
cat(latexstr)
```



# Other stuff




To get started, see the documentation '?bremla' and the associated example.

The package includes the NGRIP/GICC05 data set 'NGRIP_d18O_and_dust_5cm' downloaded from [Centre for Ice and Climate](iceandclimate.nbi.ku.dk) at the Niels Bohr Institute of the University of Copenhagen, Denmark, and the table of stadial-interstadial events presented by [Rasmussen et al. (2014)](https://www.sciencedirect.com/science/article/pii/S0277379114003485). Also includes the tie-point presented in [Adolphi et al. (2018)](https://cp.copernicus.org/articles/14/1755/2018/) and [Muscheler et al. (2020)](https://www.cambridge.org/core/journals/radiocarbon/article/testing-and-improving-the-intcal20-calibration-curve-with-independent-records/D72D9214C47FE9441B5E730D33DCCE3D). Other data and tie-points should work as well.

Warning: The provided real data example generates several thousand simulated chronologies with individual lengths exceeding 18,000. More if synchronized chronologies is also computed. This will require a significant amount of memory. If needed, reduce the number of samples generated or free up memory before use.

## Example
This is a real data example which produces 5000 simulations from a synchronized time scale.
```{r example, message=FALSE, warning=FALSE, include=TRUE}
require(INLA)
library(bremla)
data("event_intervals")
data("events_rasmussen")
data("NGRIP_5cm")

age = NGRIP_5cm$age
depth = NGRIP_5cm$depth
proxy = NGRIP_5cm$d18O

data = data.frame(age=age,dy=c(NA,diff(age)),depth=depth,depth2=depth^2,proxy=proxy)
formula = dy~-1+depth2

events=list(locations = events_rasmussen$depth,
            locations_unit="depth",degree=1)

results = bremla(formula,data,reference.label="GICC05",
                nsims=nsims,
                events=events,
                synchronization=list(method="adolphi"),
                control.fit=list(method="inla"),
                control.sim=list(synchronized=TRUE) )
summary(results)
```

```{r plot, message=FALSE, eval=TRUE, echo=FALSE}
require(ggplot2)
free_indexes = results$tie_points$free_indexes
tie_indexes = results$tie_points$tie_indexes
ageref_free = results$data$age[free_indexes]
ageref_tie = results$data$age[tie_indexes]

fullpd = data.frame(depth = results$data$depth[free_indexes],
                medians=results$simulation$summary_sync$median[free_indexes]-ageref_free,
                lower=results$simulation$summary_sync$lower[free_indexes]-ageref_free,
                upper=results$simulation$summary_sync$upper[free_indexes]-ageref_free)

gg2 = ggplot(data=fullpd,aes(x=depth)) +
geom_line(aes(y=rep(0,nrow(fullpd))),color="blue",linetype="dotted",size=0.2)+
theme_bw()+ylab("Estimated age - reference (years)")+
xlab("NGRIP depth (m)")


gg2 = gg2+geom_line(aes(y=medians))+
geom_ribbon(aes(ymin=.data$lower,ymax=.data$upper),color="red",fill="red",alpha=0.3)


gg2 = gg2 + geom_segment(data=data.frame(depth=results$data$depth[tie_indexes],                                   median=results$simulation$summary_sync$median[tie_indexes]-ageref_tie,                         lower=results$simulation$summary_sync$lower[tie_indexes]-ageref_tie,
                   upper=results$simulation$summary_sync$upper[tie_indexes]-ageref_tie),
                   aes(x=.data$depth,y=.data$lower,xend=.data$depth,yend=.data$upper),
                   col="magenta")
print(gg2)
```   

## Attribution

This code is associated and written for the papers Myrvoll-Nilsen et al. (2022) and Myrvoll-Nilsen et al. (202x) mentioned above. Feel free to use the code, but please cite the accompanying papers.

## License

The code in this repository is made available under the terms of the GNU (version >=2) License. For details, see LICENSE.md file.
